---
title: "Mixed-effects model: Dynamics of Scaffold Growth"
output:
  html_notebook: 
    fig_caption: yes
    number_sections: yes
    toc: yes
---

---

This notebook includes all analyses and plots related to the mixed-effects model described in Results section "Dynamics of Scaffold Growth."

# . Setup

```{r, message=FALSE, warning=FALSE}
if (!require(pacman)) {
  install.packages("pacman")
  require(pacman)
}

p_load(readr, dplyr, ggplot2, GGally, glmmTMB, sjPlot, DHARMa, MuMIn, patchwork,
       quantreg, effects)
```

[Back to top]("#")

---

# . Data loading

```{r, message=FALSE, warning=FALSE}
dat <- read_csv("Data/Main_DataSheet_Scaffolding_PNAS_Resubmission_FINAL.csv") %>%
  mutate(Bidirection = Bidirection / 0.5,
         Date = as.factor(Date))
dat
```

[Back to top]("#")

---

# . Exploratory Plots (not included in paper)

## . Scaffold size

### . vs angle

```{r message=FALSE}
ggplot(dat, aes(x = Angle, y = A)) +
  geom_point() + 
  geom_smooth(se = FALSE) + 
  labs(x = "Angle (degree)",
       y = "Ants in scaffold") +
  theme_light(base_size = 16)
```

### . vs traffic

```{r message=FALSE}
ggplot(dat, aes(x = Traffic_Mean, y = A, color = factor(Angle))) +
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = "Traffic (ants / sec)",
       y = "Number of ants",
       color = "Angle") +
  theme_light(base_size = 16)
```

### . vs prey transport

```{r message=FALSE}
ggplot(dat, aes(x = Prey, y = A, color = factor(Angle))) +
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = "Rate of prey transport (prey-laden ants / sec)",
       y = "Number of ants",
       color = "Angle") +
  theme_light(base_size = 16)
```

### . vs traffic bidirectionality

```{r message=FALSE}
ggplot(dat, aes(x = Bidirection, y = A, color = factor(Angle))) +
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = "Traffic bidirectionality (0 = unidirectional; 1 = bidirectional)",
       y = "Number of ants",
       color = "Angle") +
  theme_light(base_size = 16)
```

[Back to top]("#")

---

# . Data modeling

## . Scaffold size 

### . Collinearity

```{r message=FALSE}
ggpairs(data = dat, columns = c(3, 7, 8, 10)) +
  theme_light()
```

Since the analysis above revealed a strong correlation between traffic and bidirectionality, of these two, only traffic was used for the rest of the analysis. 

[Back to top]("#")

---

### . Generalized mixed-effects model with negative binomial family

#### . Model fit

```{r}
m_full <- glmmTMB(A ~ log(Angle) * Prey * Traffic_Mean + (1 | Date), data = dat, 
        family = nbinom1)
m_rank <- dredge(m_full, rank = "BIC")
m1 <- get.models(dredge(m_full, rank = "BIC"), 1)[[1]]

m_rank
m1

# warnings may appear here, which are harmless but indicate glmmTMB should be updated
```
[Back to top]("#")

---

#### . Model diagnostics

```{r}
sim <- simulateResiduals(m1, n = 1000, use.u = TRUE)
sim_dat <- tibble(scaled_resid = sim$scaledResiduals,
                  pred = sim$fittedPredictedResponse)
```

```{r}
# testUniformity() - tests if the overall distribution conforms to expectations
# testOutliers() - tests if there are more simulation outliers than expected
# testDispersion() - tests if the simulated dispersion is equal to the observed dispersion

unif <- testUniformity(sim, plot = FALSE)
disp <- testDispersion(sim, plot = FALSE)
out <- testOutliers(sim, plot = FALSE)

qq <- ggplot(sim_dat, aes(sample = scaled_resid)) +
  stat_qq(distribution = stats::qunif, alpha = 0.5) +
  stat_qq_line(distribution = stats::qunif, linetype = 2) +
  annotate("text", x = 1, y = 0, hjust = 1, vjust = -0.1,
           label = paste0("KS test: p=", format(unif$p.value, digits = 4), "\n",
                          "Dispersion test: p=", format(disp$p.value, digits = 4), "\n",
                          "Outlier test: p=", format(out$p.value, digits = 4))) +
  labs(title = "Residuals QQ plot", x = "Expected", y = "Observed") + 
  theme_light(base_size = 16)
```

```{r}
# fits a quantile regression or residuals against a predictor (default predicted value), and tests if this conforms to the expected quantile
quant <- testQuantiles(sim, plot = FALSE)

qt <- ggplot(sim_dat, aes(x = pred, y = scaled_resid)) +
  geom_point() +
  stat_quantile(size = 1, formula = y ~ x) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), linetype = 2) + 
  annotate("text", x = max(sim_dat$pred), y = 1, hjust = 1, vjust = 1,
           label = paste0("Quantile test: p=", format(quant$p.value, digits = 4))) +
  labs(title = "Residuals vs Predicted", x = "Predicted values", y = "Standardized residuals") + 
  theme_light(base_size = 12)
```

```{r}
# tests if there are more zeros than expected
zero <- testZeroInflation(sim, plot = FALSE)

zero_obs <- sum(sim$observedResponse == 0)
zero_sim <- tibble(zeros = apply(sim$simulatedResponse, 2, function(x) sum(x == 0)))

z <- ggplot(zero_sim, aes(x = zeros)) +
  geom_bar(stat = "count", alpha = 0.5) +
  geom_vline(xintercept = zero_obs, color = "red", linetype = 2, size = 1) +
  annotate("text", x = max(zero_sim$zeros), y = 150, hjust = 1, vjust = 1,
           label = paste0("Zero inflation test: p=", format(zero$p.value, digits = 4))) +
  labs(title = "Zero inflation", x = "Number of zeros in simulated values", y = "Frequency") + 
  theme_light(base_size = 12)
```
[Back to top]("#")

---

# . Final Plots

## . SI Appendix - Figure S1

```{r fig.width=4, fig.height=4}
qq / (qt | z)
#ggsave("Mixed_Effects_Model_A_Diagnostics.pdf")

```

**Model interpretation**

```{r}
summary(m1)
```

```{r}
re <- plot_model(m1, type = "re") +
  geom_hline(yintercept = 1, linetype = 2) +
  labs(y = "Relative estimates") +
  theme_light(base_size = 12) +
  theme(aspect.ratio = 2)
```

```{r}
std <- plot_model(m1, type = "std", show.values = TRUE, value.offset = .3) +
  geom_hline(yintercept = 1, linetype = 2) +
  labs(title = "Standardized effects") +
  theme_light(base_size = 12) +
  theme(aspect.ratio = 1.1)
```
[Back to top]("#")

---

## . Main Text Figure 2C

```{r fig.width=4, fig.height=2}
std + re
```

```{r}
a <- plot_model(m1, type = "eff", terms = "Angle") +
  labs(y = "Ants in scaffold", title = "Predicted counts") +
  theme_light(base_size = 12)
```

```{r}
b <- plot_model(m1, type = "eff", terms = "Prey") +
  labs(y = "Ants in scaffold", title = "Predicted counts") +
  theme_light(base_size = 12) 
```

```{r}
c <- plot_model(m1, type = "pred", 
                terms = c("Angle", "Prey [0.0000, 0.0900, 0.1580, 0.2600, 0.5450]")) +
  labs(y = "Ants in scaffold", title = "Predicted counts") +
  theme_light(base_size = 12) +
  theme(legend.position = c(0, 1),
        legend.justification = c(-0.02, 1.02),
        legend.background = element_blank())
```
[Back to top]("#")

---

## . SI Appendix Figure S2

```{r, fig.width=4, fig.height=4}
(a | b) / c 
```
