---
title: "Results: Effects of scaffolds on foraging traffic"
output:
  html_notebook:
    fig_caption: yes
    number_sections: yes
    toc: yes
---

---

This R notebook includes all code used for analysis in the Results section "Effects of scaffolds on foraging traffic." This includes the second mixed-effects model, the paired boxplot in Figure 3 (comparing the proportion of ants slipping and/or falling at the start and end of each experiment), and the SI Appendix Figures S3, S4, and S5.

# . Setup

```{r, message=FALSE, warning=FALSE}
if (!require(pacman)) {
  install.packages("pacman")
  require(pacman)
}

p_load(readr, dplyr, tidyr, broom, ggplot2, GGally, glmmTMB, MuMIn, sjPlot, viridis, DHARMa)
```

[Back to top]("#")

---

# . Data loading

```{r, message=FALSE, warning=FALSE}
dat <- read_csv("Data/Main_DataSheet_Scaffolding_PNAS_Resubmission_FINAL.csv") %>%
  select(., Date, Angle, Start_Traffic, End_Traffic, Prey, Bidirection, SlipFalls_Start, SlipFalls_End) %>%
  mutate(Start = (SlipFalls_Start / 120) / Start_Traffic,
         End = (SlipFalls_End / 120) / End_Traffic) %>%
  pivot_longer(., cols = c("Start", "End"), names_to = "Time", values_to = "SlipFalls_Prop") %>%
  mutate(., Time = factor(Time, levels = c("Start", "End")),
         W = Start_Traffic * 120)

dat
```

[Back to top]("#")

---

# . Slipping comparisons

## . Figure 3: Paired boxplot for all angles

Estimating effects of scaffold formation on traffic at each angle. Comparing the proportion of ants slipping/falling at experiment start (first 2 minutes, shown in dark green) and end (final 2 minutes, shown in light green), representing the absence or presence of scaffold structures, respectively.

```{r message=FALSE}
ggplot(dat, aes(x = factor(Angle), y = SlipFalls_Prop, color = Time, fill = Time))  + 
  scale_color_viridis(discrete = TRUE, direction = 1, begin = .3, end = .7) +
  scale_fill_viridis(discrete = TRUE, direction = 1, begin = .59, end = .96) +
  geom_point(pch = 20, alpha = .8, position = position_jitterdodge(dodge.width = .7)) +
  geom_boxplot(width = .69, alpha = 0.58, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  theme_bw(base_size = 18) +
  #ylim(0, 0.06) +
  theme(aspect.ratio = .66, axis.text = element_text(size = 12), legend.position = c(0, 1),
        legend.justification = c(-0.1, 1.1), legend.title = element_blank(),
        #panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5)) +
  xlab("\nSurface angle") + ylab("Proportion of ants \n slipping and/or falling")
 #ggsave("Figure3_FINAL_Resubmission2.pdf")
 

```

[Back to top]("#")

---

# . Data modeling

## . Collinearity

```{r message=FALSE}
ggpairs(data = dat, columns = c(2, 4)) +
  theme_light()
```

### . Generalized mixed-effects model with gaussian family

**Model fit**

```{r}
m_full <- glmmTMB(SlipFalls_Prop ~ Time * Angle + (1 | Date), 
                  data = dat, family = binomial, weights = dat$W,
                  na.action = na.fail)
m1 <- get.models(dredge(m_full, rank = "BIC"), 1)[[1]]
m1
summary(m1)
get_model_data(m1)


```

**Model diagnostics**

```{r fig.width=2, fig.height=2}
plot_model(m1, type = "diag", terms = "Date")$Date +
  theme_light(base_size = 12)
```

**Model interpretation**

```{r}
summary(m1)
plot_model(m1)
```

```{r}
re <- plot_model(m1, type = "re") +
  geom_hline(yintercept = 1, linetype = 2) +
  labs(y = "Relative estimates") +
  theme_light(base_size = 12)
```

```{r}
std <- plot_model(m1, type = "std") +
  geom_hline(yintercept = 1, linetype = 2) +
  labs(title = "Standardized effects") +
  theme_light(base_size = 12)
```

```{r fig.width=4, fig.height=2}
std + re
```


```{r}
plot_model(m1, type = "eff", terms = "Angle") +
  ylim(0, NA) +
  labs(y = "Proportion of ants \n slipping and/or falling", title = "Predicted proportion") +
  theme_light(base_size = 16)
```

```{r}
plot_model(m1, type = "eff", terms = "Time", show.data = FALSE) +
  ylim(0, NA) +
  labs(y = "Proportion of ants \n slipping and/or falling", title = "Predicted proportion") +
  theme_light(base_size = 16)
```

```{r fig.width=2, fig.height=2}
plot_model(m1, type = "eff", terms = c("Angle", "Time"), show.data = TRUE) +
  ylim(0, NA) +
  labs(y = "Proportion of ants \n slipping and/or falling", title = "Predicted proportion") +
  theme_light(base_size = 12) +
  theme(legend.position = c(0, 1),
        legend.justification = c(-0.02, 1.02),
        legend.background = element_blank())
```

